---
title: "Testing models of secondary production: simple to advanced"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Testing models of secondary production: simple to advanced}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
- \usepackage{lineno}
- \usepackage{amsmath}
- \numberwithin{equation}
bibliography: refs.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE, 
  comment = "#>"
)
```

```{r setup}
library(secpRod)

load('data/singleCohortSim.rda')

```

# Purpose

This document walks through the different methods and models to calculate secondary production. We start with a single simulated univoltine population for simplicity. The details for this simulation can be found in [Simulating the sampling of populations](https://jimjunker1.github.io/secpRod/articles/sampling-simulation.html) vignette. 

## Introduction 

Currently, the analysis is fully implemented for the size-frequency and production:biomass ratio methods. This quick tutorial will outline steps to calculate secondary production for populations and whole communities. First, load the package with `library(secpRod)`.

The package comes with a number of data sets of macroinvertebrate species and community data:

1) simulated data set of a single univoltine species accessed with `data("singleCohortSim")`

2) a full community data set from @junker2014 that can be accessed with `data("wbtData")`

The simulated data set is a single data frame of artificially sampled density and size-frequency data through time. This 

```{r echo = TRUE}
data("singleCohortSim")

#clean this data set to align 
head(singleCohortSim)
```



# References

